---
title: "p8105_hw3_csl2209"
author: "Casandra Laney"
date: "2025-10-07"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Problem 1

Load the data

```{r}

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

library(lubridate)
library(patchwork)
library(p8105.datasets)
data("instacart")

```

The dataset 'instacart' contains 15 variables of grocery data collected through the instacart ordering service from June 24, 2017. 

Important variables include:

* 'order_id': the ID number of the order.

* 'product_id': the ID number of the product ordered.

* 'add_to_cart_order': the order by which the product was added to the cart

* 'reordered': whether the customer has ordered the item previously or not

* 'user_id': the identification number of the user who ordered.

For example, in the first row of data, we see that order 1 contained the product identified as 49302 one time. This product was reordered once by user 112108.

```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_obs = n()) |>
  arrange(desc(n_obs))
```

There are 134 aisles, and the most items are ordered from the 'fresh vegetables' aisle.

```{r aisle_plot, fig.width=10, fig.height=6}
instacart |>
  group_by(aisle) |>
  summarize(n_obs = n()) |>
  filter(n_obs > 10000) |>
  arrange(n_obs) |>
  ggplot(aes(x = reorder(aisle, n_obs), y = n_obs)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Items Ordered by Aisle (>10,000 Orders)",
    x = "Aisle",
    y = "Number of Items Ordered",
    caption = "Data from the Instacart dataset"
  ) 
```

There are 134 aisles, and the most items are ordered from the 'fresh vegetables' aisle.

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  filter(
    aisle == "baking ingredients" |
    aisle == "dog food care" |
    aisle == "packaged vegetables fruits"
  ) |>
  group_by(aisle, product_name) |>
  summarize(n_orders = n()) |>
  arrange(aisle, desc(n_orders)) |>
  mutate(rank = row_number()) |>
  filter(rank <= 3) 
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )

```

# Problem 2

 Import, clean, and otherwise tidy these datasets.
 
```{r}

zori = read_csv("data/zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") |>
  janitor::clean_names() |>
  drop_na() |>
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "date", 
    values_to = "zori"
    ) |>
    mutate(
      date = str_remove(date, "^x"),
      zip_code = as.factor(region_name)) |>
  select(-region_name, -region_type, -state, -city, -state_name, -county_name, -metro)
  

zip_codes = read_csv("data/zillow_data/Zip Codes.csv") |>
  janitor::clean_names() |>
  mutate(
    zip_code = as.factor(zip_code),
    neighborhood = as.factor(neighborhood),
    ) |>
  select(-file_date)

tidy_zillow =
  zori |>
  left_join(zip_codes, by = "zip_code") |>
  mutate(
    year = year(ymd(date))) |>
  select(date, zip_code, neighborhood, county, zori, state_fips, county_code, county_fips, region_id, size_rank)

```
 
 
 There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?
 
```{r}
zip_count =
  tidy_zillow |>
  group_by(zip_code) |>
  summarize(n_months = n())

high_count =
  zip_count |>
  filter(n_months == 116)

low_count =
  zip_count |>
  filter(n_months < 10)

```
 
 
 Create a reader-friendly table showing the average rental price in each borough and year (not month). Comment on trends in this table.
 
```{r}

tidy_zillow |>
  mutate(date = ymd(date),year = year(date)) |>
  group_by(
    county, year) |>
  summarize(
    avg_rent = mean(zori)) |>
  pivot_wider(
    names_from = year, values_from = avg_rent) |>
  arrange(county)

```
 
 
 Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.

```{r zillow_plot, fig.width=11, fig.height=6}
all_years = 
  tidy_zillow |>
  mutate(date = ymd(date)) |>
  ggplot(aes(x = date, y = zori, group = zip_code, color = county)) +
  geom_line() +
  labs(
    title = "NYC Rental Prices by Zip Code (2015-2024))",
    x = "Date",
    y = "Rental Price",
    color = "Borough"
  )  +
  scale_x_date(
    date_breaks = "1 year")
```


Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot.

```{r}
average_price =
  tidy_zillow |>
  mutate(date = ymd(date)) |>
  filter(year(date) == 2023) |>
  group_by(zip_code, county, date) |>
  summarize(avg_zori = mean(zori)) |>
  ggplot(aes(x = date, y = avg_zori, group = zip_code, color = county)) +
  geom_line() +
  labs(
    title = "NYC Rental Prices by Zip Code (2023))",
    x = "Month",
    y = "Rental Price",
    color = "Borough"
  )  +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%m")
```

Combine the two previous plots into a single graphic, and export this to a results folder in your repository.

```{r combined_plot, fig.width=11, fig.height=6}

combined_plot =
  all_years / average_price + plot_layout(guides = "collect")

ggsave(
  filename = "results/nyc_rental_combined.png",
  plot = combined_plot,
  width = 12,
  height = 10,
  dpi = 300
)
```

# Problem 3

Load, tidy, merge, and otherwise organize the data sets. 

```{r}
accel = read_csv("data/nhanes_accel.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  drop_na() # exclude missing
  
covar = read_csv("data/nhanes_covar.csv", na = c("NA", ".", ""), skip = 4) |>
  janitor::clean_names() |>
  drop_na() # exclude missing

accel_combined = 
  accel |>
  left_join(covar, by = "seqn") |> # merge datasets
  select(seqn, sex, age, bmi, education, everything()) |>
  filter(age >= 21) |> # exclude participants less than 21
  mutate( 
    seqn = as.factor(seqn),
    sex = 
      case_match(
        sex,
      1 ~ "male",
      2 ~ "female"
    ),
    education =
      case_match(
        education,
        1 ~ "Less than High School",
        2 ~ "High School Equivalent",
        3 ~ "More than High School"
      ))
```

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}

accel_combined |>
  filter(sex == c("male", "female")) |>
  group_by(education, sex) |>
  summarize(count = n()) |>
  pivot_wider(names_from = sex, values_from = count)

accel_combined |>
  filter(sex == c("male", "female")) |>
  ggplot(aes(x = age, fill = sex)) +
  geom_histogram(position = "dodge", binwidth = 10, alpha = 1) +
  facet_wrap(~education) +
  labs(
    title = "Age Distribution by Sex and Education",
    x = "Age",
    y = "Count",
    fill = "Sex")
    
```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}


```

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}

```

