---
title: "p8105_hw3_csl2209"
author: "Casandra Laney"
date: "2025-10-07"
output: github_document
---


```{r}
library(tidyverse)

```

## Problem 1

Load the data

```{r}
library(p8105.datasets)
data("instacart")

```

The dataset 'instacart' contains 15 variables of grocery data collected through the instacart ordering service from June 24, 2017. 

Important variables include:

* 'order_id': the ID number of the order.

* 'product_id': the ID number of the product ordered.

* 'add_to_cart_order': the order by which the product was added to the cart

* 'reordered': whether the customer has ordered the item previously or not

* 'user_id': the identification number of the user who ordered.

For example, in the first row of data, we see that order 1 contained the product identified as 49302 one time. This product was reordered once by user 112108.


How many aisles are there, and which aisles are the most items ordered from?'

```{r}

```


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}

```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}

```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}

```

# Problem 2

 Import, clean, and otherwise tidy these datasets.
 
```{r}


```
 
 
 There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?
 
```{r}

```
 
 
 Create a reader-friendly table showing the average rental price in each borough and year (not month). Comment on trends in this table.
 
```{r}

```
 
 
 Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.

```{r}

```

Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot.

```{r}

```

Combine the two previous plots into a single graphic, and export this to a results folder in your repository.

```{r}

```

# Problem 3
